\documentclass[a4paper,11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{hyperref}
\usepackage{authblk}

\title{A Statically Checked Interface and a Checker for Tezos Smart Contracts}
\author{Peter Thiemann}
\author{Thi Thu Ha Doan}
\affil{University of Freiburg, Germany}
\begin{document}
\maketitle{}

\section{Report January-June 2023}
\label{sec:executive-summary}

We proposed the following activities and deliverables for the
project. We started all planned activities for the reporting period
and we already started some planned activities for the next period.

Currently, the project is on track and making good progress, but Ms.\ Doan is on maternal leave
and it is unclear if she will take parental leave thereafter. If she
does, it may lead to unforseeable delays.

\subsection{January 2023 - June 2023}

\begin{itemize}
\item Proposed activity: Development of a dynamic logic as a formal foundation for symbolic interpretation of Michelson. \\
  $\Rightarrow$ This activity progressed significantly with the help of a master's
  student supervised by Thiemann. The activity started later than
  planned because the student was unavailable till the end of the
  teaching period in March.
\item Deliverable: Formalization (in Agda) of a dynamic logic for a core fragment of Michelson with soundness proof. \\
  $\Rightarrow$ This deliverable is currently in beta and will be finished when the
  master's thesis submission is due in September 2023.
\item Deliverable: Paper about this formalization (writeup starting
  July 2023). \\
  $\Rightarrow$ The paper will be based on the master's thesis, so the final writing
  starts after the submission in September. The goal is to submit to a
  conference during the winter/early spring.
\end{itemize}

\subsection{July 2023 - December 2023}
\label{sec:july-2023-december}

\begin{itemize}
\item Activity: Implementation of symbolic interpreter for a core
  fragment of Michelson in OCaml based on an SMT solver, most likely
  on Z3. \\
  $\Rightarrow$ This activity has been driven by Doan between January and June
  2023. 
\item Deliverable: Symbolic interpreter. \\
  $\Rightarrow$ The current prototype of the symbolic interpreter is available here
  \url{https://github.com/proglang/tezos-project/tree/master/static_checker}
\item Deliverable: Case studies with example contracts.
\item Deliverable: Paper or tutorial on this tool (writeup starting
  January 2024) \\
  $\Rightarrow$ A draft is available here \url{https://github.com/proglang/tezos-project/tree/master/symbolic-execution/paper}
\end{itemize}

\subsection{January 2024 - June 2024}
\label{sec:january-2024-june}

Not started.

\begin{itemize}
\item Activity: Design and implementation of a DSL for static contract properties. This DSL will compile to SMT and interface with the symbolic interpreter to perform verification.  
\item Deliverable: Implementation of the DSL aka contract checker.
\item Deliverable: Report on this tool.
\end{itemize}

\subsection{July 2024 - December 2024}
\label{sec:july-2024-december}

Not started.

\begin{itemize}
\item Activity: Complete design and implementation of the contract module DSL including contract orchestration. 
\item Deliverable: Full implementation of contract modules.
\item Deliverable: Followup on our FMBC paper. 
\end{itemize}

\bibliography{report}
\bibliographystyle{plainurl}
\end{document}
