\section{Advice}\label{sec:advice}

This section gives some advice how to write a thesis ranging from writing style to formatting. To be sure, ask your advisor about his/her preferences.\\
For a more complete list we recommend to read Donald Knuth's paper on mathematical writing. (At least the first paragraph). \url{http://jmlr.csail.mit.edu/reviewing-papers/knuth_mathematical_writing.pdf}

    \begin{itemize}

        \item If you use formulae pay close attention to be consistent throughout the thesis!

        \item In a thesis you never write `In [24] the data is..'. You have more space than in a paper, so write `AuthorXY et al. prepare the data... [24]'. The rule is that the text should remain grammatically correct even if the reference is erased.
          For that reason, the template uses \texttt{natbib} with style \texttt{plainnat}, which makes it much easier to stick to the rule. Most of the time the citation should be at the end of the sentence before the full stop with a no-break space. \verb|... last word~\cite{XY}.| With \texttt{natbib} you can use the \texttt{citet} command to refer to the authors' name and cite at the same time. See \url{https://www.sharelatex.com/learn/Bibliography_management_with_natbib} for more information.
          
        \item Get your bibtex entries from \url{https://dblp.org/}.

        \item Pay attention to comma usage, there is a big difference between English and German. `...the fact that bla...' etc.

        \item Do not write contractions like `don't ', `can't', and so on. Instead write `do not', `cannot', and so on.

        \item If an equation is at the end of a sentence, add a full stop. If it is not the end, add a comma: {$a= b + c$~~~~(1),}

        \item Avoid footnotes if possible.

        \item Use \verb|``''| for citing, not \verb|""|.
        \item Captions of tables and figures do not end with a period, unless the caption is a full sentence.
        \item Titles and headings ought to be capitalized properly. Here is a tool to help: \url{https://capitalizemytitle.com/}

        \item Check the spelling of your thesis before submitting. There are tools like \texttt{aspell} that help you find such mistakes. 
        As a spell checker cannot find grammatical errors, it does not relieve you from properly reading your thesis again and from getting comments from somebody else.
        You can find an introduction under \url{https://git.fachschaft.tf/fachschaft/aspell}.

      \item If the thesis contains graphs or other drawings consider
        using \texttt{tikz}. Do not waste you time studying its manual
        except for the introduction and some examples, instead go for
        some of the examples available on the net and adapt. For
        function graphs or diagrams consider using \texttt{pgfplots}
        or \texttt{gnuplot}.  Using latex-related tools has the
        advantage that the style is more consistent (same font,
        formatting options etc.) than with some external program.

      \item Discuss with your advisor whether to use passive voice or
        not. In most computer science papers passive voice is
        avoided. It is harder to read, more likely to produce errors,
        and most of the times less precise. Of course, there are
        situations where passive voice fits but in scientific papers
        they are rare. Compare the sentence: `We created the wheel to
        solve this.' to `The wheel was created to solve this'.  You
        don't know who did it, making it harder to understand what is
        your contribution and what is not.

        The thesis should not be written in first person singular. First person plural is generally accepted, but it is good practice to avoid the pronouns `we', `our' etc as much as possible.

    \end{itemize}