\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{mathpartir}
\usepackage{listings}
\usepackage{url}

\usepackage{xcolor}
\usepackage{tcolorbox}
\tcbuselibrary{breakable}

\newenvironment{changethis}{%
  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=blue!20,colframe=blue!20]}{%
  \end{tcolorbox}}

\newenvironment{ethereum}{%
  \begin{tcolorbox}[breakable,notitle,boxrule=0pt,colback=red!20,colframe=red!20]}{%
  \end{tcolorbox}}


\input{solidity-highlighting}

\title{Expos\'{e} Master Thesis\\
  Checking Assertions for Smart Contracts}
\author{Peter Thiemann}
\begin{document}
\maketitle{}

\section{Introduction}
\label{sec:introduction}

Every computation performed by a Smart Contract on the blockchain generates costs. Each
unit of computation and each unit of storage used by an algorithm must be paid for. To
avoid this cost, an application might perform some computation away from the blockchain
(i.e., off-chain) and submit the result as a parameter to a contract on the
blockchain. Typically, such a computation asserts certain properties of the 
submitted parameter. 

However, this approach raises the issue that on the one hand the contract should take
advantage of the offchain computation and assume that the submitted parameter has
these properties, but on the other hand, the off-chain computation might be wrong and
submit illegal parameters. So, we need a mechanism that checks the validity of the
assumptions before the contract starts executing.

As an example, consider a contract that takes a prime number as a parameter.
\begin{lstlisting}[numbers=none]
contract Example {
  function (int p) public {
    // assume p is prime
    ...
  }
}
\end{lstlisting}
This assumption can be expressed with an explicit assertion in predicate logic.
\begin{displaymath}
  (\forall n) (2 \le n \le \sqrt p) \Rightarrow (p \mathbin{\%} n) \ne 0
\end{displaymath}
To test the validity of this assumption requires a loop in the contract, but the test would take
$O(\sqrt p)$ time (assuming constant time for computing the remainder) and thus produce extra cost
linear in $\sqrt p$ accordingly. 
However, we could do better by recruiting the validators of the contract for a distributed
effort to find a counterexample. To this end, we consider the negation of the assertion.
\begin{gather}\label{eq:4}
  (\exists n) (2 \le n \le \sqrt p) \wedge (p \mathbin{\%} n) = 0
\end{gather}
This assertion can be checked pointwise by having each validator independently choose a
random $n$ fulfilling $2 \le n \le 
\sqrt p$ and checking whether $(p \mathbin{\%} n) = 0$. If the remainder is $0$, the
validator found a counterexample, posts its veto to the P2P net, and stops further
exection. Otherwise, it accepts $p$ knowing that other points will be checked by other
validators.

In this scenario, each validator only needs to be paid to generate a random number and
perform a division, which is a constant cost independent from $p$.

Of course, this validation is only probabilistic, so its effectiveness depends on the
number of validators. One could say that the community of validators implements a Bloom
filter for the set of primes: if a value $p$ is rejected it is certainly not a prime (because
there exists a counterexample); if a value $p$ is not rejected it is prime with a
probability that depends on $p$ and the number of validators. 

As another example, consider a contract that takes a sorted array of integers.
\begin{lstlisting}[numbers=none]
contract Sorted {
  function find (int[50] a, int v) public {
    // assume a is sorted
  }
}
\end{lstlisting}
The explicit assertion would be
\begin{gather}\label{eq:1}
  (\forall k) (0\le k <49) \Rightarrow a[k] \le a[k+1]
\end{gather}
While we can check this contract in $O(1)$ time, the constant factor is big! So we
consider its negation.
\begin{displaymath}
  (\exists k) (0\le k <49) \wedge a[k] > a[k+1]
\end{displaymath}
Again, we can have every validator generate a random number $k$. If the
condition is true for such $k$, then the validator found a counterexample for the sortedness
of the array. Otherwise, the validator relies on the other validators to check
different numbers.

To obtain an estimate of the number of validators needed to find a counterexample with
high probibility, 
let's assume the array is unsorted only at position $0$, the size of the array is $n$,
and the number of validators is $m$. Each validator independently has a probability of
$1/n$ to detect the problem and thus probability $\frac{n-1} n$ not to detect the
problem. Hence, if we assume that each $k$ is chosen independenty from a uniform distribution,
the probability that no validator checks at
position $0$ converges to $0$ as the number of validators approaches infinity.
\begin{displaymath}
  \lim_{m\to\infty}\frac{(n-1)^m}{n^m}
  = \lim_{m\to\infty} \left(\frac{n-1}{n}\right)^m
  = 0
\end{displaymath}

In Dafny (citation) \url{https://rise4fun.com/Dafny/tutorialcontent/guide#h29} you can
write
\begin{lstlisting}
forall j, k :: 0 <= j < k < a.Length ==> a[j] <= a[k]
\end{lstlisting}
to express that an array is sorted. This predicate is equivalent to the one given in
\eqref{eq:1}, but it might be more challenging to test. Its negation is
\begin{gather}
  \label{eq:2}
  (\exists j, k ) (0\le j< k < |a|) \wedge a[j] > a[k]
\end{gather}
So we'd have to generate two random numbers $j$ and $k$ such that the condition $0 \le
j < k < |a|$ is fulfilled.


Another condition that might be tested on an array is the heap condition
\begin{gather}
  \label{eq:3}
  (\forall i) (0 \le i < \lfloor|a|/2\rfloor) \Rightarrow a[i] \le a[2i+1] \wedge (2i+2
  < |a| \Rightarrow a[i] \le a[2i+2])
\end{gather}


\section{Tasks}
\label{sec:tasks}

\subsection{Offline Design}
\label{sec:offline-design}

\begin{itemize}
\item Formally define a set of logical formulas that are amenable to the kind of checking
  described above. In a first approximation, the interesting assertion formulas seem to
  be those with prenex universal quantification. 
\item Collect examples for interesting formulas.
\item Formalize the cost incurred by checking a formula.
\item Extend the contract language with assertions stating these logical formulas.
\item Implement the transformation: negation of the formula followed by transformation
  of existentials to picking a random value. For initial experiments, the target of the
  translation can be any language; later on, the target has to be a language hosted on
  the blockchain.

  Negation turns prenex universial quantification into prenex existential quantification. But it
  would be ineffective to generate a random integer and then check that it falls in the
  interesting range (e.g., $2\le n \le \sqrt p$ and $0\le k< n-1$). Rather, the formula
  needs to be analzed and the random generator should be restricted to only generate
  interesting values. 

  The random choice should be performed in a ``clever'' way that does not waste too
  much effort. For a wasteful example, consider~\eqref{eq:2}:
  \begin{gather*}
    (\exists j, k ) (0\le j< k < |a|) \wedge a[j] > a[k]
  \end{gather*}
  In the worst case, we would generate two arbitrary non-negative integers $j$ and $k$
  and then check the predicate. However, the predicate $(0\le j< k < |a|)$ will fail
  almost always, so we do not get to test the interesting part of the predicate
  $a[j]>a[k]$.

  In the best case, the generated testing code would be equivalent to the following code fragment
  in Python
\begin{lstlisting}[language=Python]
def exists_counterexample (a : list) -> bool:
  if len(a) < 2: return False
  j = random.randint (0, len(a) - 2)
  k = random.randint (j+1, len(a) - 1)
  return a[j] > a[k]
\end{lstlisting}

  As another, simpler example consider~\eqref{eq:4} (checking for prime numbers)
\begin{lstlisting}[language=Python]
def exists_counterexample (p : int) -> bool:
  r = math.isqrt (p)
  if r < 2: return False
  n = random.randint (2, r)
  return (p % n) == 0
\end{lstlisting}

  Observe that in each case there is a test to eliminate trivial cases, which avoids
  illegal invocations of \lstinline/randint/.
\item Each test function should come with a formula that gives an estimate of
  its effectiveness, i.e., how many times do we have to run the test to find a
  counterexample with a given probability $p$. Ideally, the formula returns a lower
  bound on the number of tests for any $p$.
\item What is the minimum number $m$ of validators such that the probability of
  spotting an error is less than a given threshold $q$?
\item Instead of using random numbers, is there a way to coordinate a systematic
  exploration of the iteration space? (Is this at all necessary?)
\item What are the connections to random testing? 
\item I think this scheme implements a Bloom filter, globally? We are testing whether a value is a
  member of an implicitly given set. We may accept a value even if it does not belong
  to the set, but if we reject a value, it definitely does not belong to the set.
\end{itemize}

\subsection{Blockchain-Specific Tasks}
\label{sec:blockch-spec-tasks}

\begin{ethereum}
  Ethereum-specific remarks:
  Implementing this scheme requires adding instructions to the virtual machine. One
  instruction would generate a random number (EVM uses 512 bit unsigned integer IIRC),
  the other would raise an overriding failure. This failure will have to indicate that
  validation of the contract fails.
\end{ethereum}

\begin{itemize}
\item The VM interpreter needs to be extended with these instructions.
\item The translation (of formulas) needs to retargetted to VM code. This may have to
  be hacked into the compiler.
\item The translation has to be amended so that counterexamples can also be checked.
\item (extension) Every validator could perform a predetermined finite number of checks for each
  assertion. This would not be significantly more expensive. Moreover, the number of
  checks could be set individually depending on the contract and on the size of the
  input.
\item (extension) One could also check properties of the values stored with the contract.
\end{itemize}
\subsection{Protocol Design}
\label{sec:protocol-design}

\begin{changethis}
  Tezos specific remarks: here the number of validators is really small compared to
  Bitcoin or Ethereum
  \begin{itemize}
  \item Maybe we need to propose and validate the same block multiple times before it
    gets accepted. E.g, a transaction containing this feature is only accepted if it is
    revalidated $n$ times.
  \item Extra validators only for this kind of transaction.
  \item Maybe each validator needs to conduct multiple test runs.
  \item Maybe test runs need to be coordinated
  \end{itemize}
\end{changethis}

\begin{itemize}
\item   Once a counterexample has been found,
  this counterexample is broadcast to the network where it can be verified by further
  nodes.  I think this broadcast and the reaction to this broadcast requires an update of
  the protocol: all receivers of the broadcast will check the counterexample and revert
  the transaction. 
\item A transaction using counterexamples has to
  wait for a couple of cycles in the transaction pool until the node feels safe to
  include the transaction in a block.
  The number of cycles to wait depends on the contract and on the size of the input.
  So that number should be variable.
\item (low priority) This approach poses the danger of a DOS attack. Someone could post a false
  counterexample, which does not really lead to a failure. If everyone switches to
  check this proposed counterexample instead of generating a random number, then the
  real problem may remain hidden. So, there has to be a component in the protocol that
  penalizes someone who broadcasts a false counterexample. Moreover, not all validators
  should switch to recheck a proposed counterexample; or they could check it in
  addition to their own random check.

  More precisely:
  \begin{itemize}
  \item Every validator must run at least one (round of) random tests.  
  \item If it receives a proposed counterexample, it must check the counterexample,
    too.
  \item It remains to investigate a countermeasure for nodes generating a flood of
    false counterexamples, which would keep every validator busy. Perhaps checking of
    the counterexample should also be done probabilistically.
  \end{itemize}
\end{itemize}

\section{Rejected Ideas}
\label{sec:rejected-ideas}

This section contains some ideas that we discussed but rejected. Feel free to resurrect
any of this if you find new arguments.

\begin{itemize}
\item Q: How about instead of generating and checking random number, it could
  store a Bloom filter for already checked number, so a validator can
  generate a random number and check it if it is rejected by Bloom filter,
  but this also happen that there will be some number never get checked
  because it is rejected by Bloom filter, but actually never been checked.

  A: This is not a good idea for several reasons.
  \begin{enumerate}
  \item It's not clear where we should store the Bloom filter. It has to be available
    everywhere, so it has to be stored on the blockchain.
  \item If it's on the blockchain, then it serializes the tests for counterexamples:
    the point of the scheme is to test many random values concurrently without
    interference.
  \item The Bloom filter errs in the wrong direction. It may accept elements that have
    not been added before, which means that elements that should be tested are not. We
    would need a co-Bloom filter that errs in the other direction: it should reject all
    elements that have not been added, and it may also reject elements that have been added.
  \end{enumerate}

\end{itemize}

\end{document}

